How wide spread deflections bias brain-wide models of resting state fMRI.

### 2020-02-22 manuscript/manuscript.tex
some updates... still needs a lot of work!

The results are structured as follows. Firstly, we investigate the spatiotemporal characteristics of the data that need to be modelled. Secondly, we then investigate the spatiotemporal characteristics of the models themselves. Thirdly, we describe the accuracy of the fits to data. With these results, we then employ heterogenous models to see if they help.

@@

Firstly, an important technique to analyze the quality of resting state fMRI scans is to visualize the spatiotemporal time series of scans. As shown in Figure~\ref{fig:rsfMRIexp}C, (MPP looks pretty similar, might include a noisier subject..) that following minimal preprocessing (MPP) a ``low-motion'' subject from the UCLA cohort (sub-10274) contains a wealth of wide-spread signals that are times tied to motion events as captured by the mean framewise displacement (FD) (in the blue rectangles). Additional wide-spread signals have also been tied to physiological events such as respriation \cite{power2017simple,power2018ridding} or heart rate variability \cite{birn2006separating}. A robust de-noising method - ICA-AROMA with regression of white and CSF signals (+2P) mitigates some of these events but as often with these de-noising methods WSDs remain \cite{power2018ridding,power2013steps} as can be appreciated in Fig.~\ref{fig:rsfMRIexp}D. The link between these WSDs and physiology and motion thus treats these particular WSDs as noise, and can be appreciated in the functional connectivity matrix in Fig~\ref{fig:rsfMRIexp}G, where the globally coherent structure causes a FC matrix that is postively correlated. This is an important issue as that if the FC matrix is used to compare between groups, then large motion drives the changes in the edges of the FC matrix rather than underlying neuronal processes. For these reasons, further de-noising is often employed to rid rsfMRI of these artefactual signals. As described in the methods GMR is regularly used to rid rsfMRI of these influences, and as seen in Fig~\ref{fig:rsfMRIexp}E, it can help mitigate these effects however introducing negative correlations due to this zero-centering \cite{murphy2009impact,saad2012-gsr-neg} which can be apprecaited in Fig~\ref{fig:rsfMRIexp}H and Fig\ref{fig:rsfMRIexpbiphasic}H. Using the mean of all voxels to remove WSDs can be problematic if the WSD is made of components that exist in multiple phases such as the subject in Fig.\ref{fig:rsfMRIexpbiphasic} where the common underlying WSD is expressed in two clear phases - i.e. the top half has the opposite sign of the bottom half. As can be apprecated in Fig.\ref{fig:rsfMRIexpbiphasic}D vs Fig.\ref{fig:rsfMRIexpbiphasic}E, GMR does not remove this WSD adequately, leaving behind a spatiotemporal time series that still contains artefactal WSD. This characterization of the types of WSDs is explained in \cite{aquino2019identifying} and can be thought to be of three types: weak (not shown here), single-phase, and biphasic. 
\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/figure_example.png}
\caption{Example carpet plots, single phase.}\label{fig:rsfMRIexp}
\end{figure*}

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/figure_example2.png}
\caption{Example carpet plots noisy biphasic subject.}\label{fig:rsfMRIexpbiphasic}
\end{figure*}

For these reasons, a new techinque - DiCER - was developed to targets these weak diffuse signals\cite{aquino2019identifying}. On application of DiCER to the ICA-AROMA+2P output form all subjects, the carpet plots in Figures \ref{fig:rsfMRIexp}F and \ref{fig:rsfMRIexpbiphasic}F show the reduction of WSDs, whilsting revealing FC matrices in Figures \ref{fig:rsfMRIexp}I and \ref{fig:rsfMRIexpbiphasic}I that have structure without zero centring the functional connectivity matrix. We note that undergoing network discovery, DiCER appears to improve statistical sensivity of neuronal components - or at worst - on par with GMR \cite{aquino2019identifying}. These type of WSDs are present across the three datasets described, and the removal of the WSDs by DiCER or GMR generally improves QC-FC measures (not sure how much detail...).

@@

The above results illustrate that the nature of coherent structures are largely seen as artefactual and the presence of these structures removed with varying degrees of de-noising. The varying degress of WSDs are important when biophysical models of resting state fMRI, as the models not only have to capture the overall functional connectivity but the typical dynamics within data. As described in the methods, a key feature of all biophysical models is the global coupling parameter $G$, that changes the dynamics of the models from un-coupled dynamical nodes ($G=0$) through to completely synchronized dynamics ($G>4$ in the models considered here). Therefore, the tunining of the model and the dynamic regime that is being fit will depend on the quality of the data wiht regards to the level of WSDs. This can be appreciated by analzying the WSDs in the biophysical predictions in a similar fashion to the experimental data.

Firstly, for the Balanced Excitation Inihbition model Fig.~\ref{fig:BalancedEI_G}~A shows that varying the global coupling parameter $G$ largely changes the dynamics of the model. If we compare the carpet plots of varying levels of the coupling parameter $G$ we see visually that the level of WSDs peaks for $G=3$, and is reflected in the global structure in the functional connectivity matrices in Fig.~\ref{fig:BalancedEI_G}~B. The level of WSDs can be quantified by using heuristics, namely using the variance explained by the first PC (VE1) or the global mean of all correlations in the FC matrix (GFC). As shown in Figs~\ref{fig:BalancedEI_G}~C, and ~\ref{fig:BalancedEI_G}~D we can see that like in the carpet plots and the FC matrix, VE1 and GCF peaks for $G=3$ meaning that this parameter maximises the coherence and thus WSDs. When comparing these measures to subject 10274 in the UCLA data we find that the level of synchrony needed to match the data depends on the level of WSDs within the data that are dependendent on the preprocessing pipeline. Overall, these features are shown in the neural mass model in Fig.~\ref{fig:BalancedEI_G}~F--I as well as the Hopf bifurcation model in Fig.~\ref{fig:BalancedEI_G}J--M. Where the transition from uncoupled noisy nodes changes from obtaining an optiumum in the Balanced EI case (Fig.~\ref{fig:BalancedEI_G}C,D), to a phase transition in the Neural mass model (Fig.~~\ref{fig:BalancedEI_G}~H,I) to a gradual increase in coherence in the Hopf bifurcation model(~\ref{fig:BalancedEI_G}~L,M).

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/Models_all.png}
\caption{Wide spread deflections in the Balanced Excitation-inhibiton model. A B C D}\label{fig:BalancedEI_G}
\end{figure*}

% \begin{figure*}[ht!]
% \includegraphics[width=1\textwidth]{figs/BTFImage.png}
% \caption{Wide spread deflections in the neural mass model. A B C D}\label{fig:NM_G}
% \end{figure*}

% \begin{figure*}[ht!]
% \includegraphics[width=1\textwidth]{figs/HopfImage.png}
% \caption{Wide spread deflections in the Hopf Bifurcation model. A B C D}\label{fig:Hopf_G}
% \end{figure*}

The above results indciate that data with WSDs requires models to be biased toward the synchronized regime, whereas data with reduced WSDs push data to the unsynchronized regime (i.e. lower $G$). We can explore this issue further in the highly synchronized regime, and the weakly synchronized regime. Firstly, models within the synchronized regime appear to have a globaly coherent signal is present in the spatiotemporal carpet plots as visually expressed by the large vertical stripes in Figs~\ref{fig:BalancedEI_G}~A,F,J with $G=3,3.25,2--8$ respecively. If we were to model this behaviour explicitly, we first note that in all three models the local dynamics at node $i$ are modulated by the influences owing to the sum (or mean) of the neighbouring nodes weighted by $C$ i.e. by a general factor
\begin{equation}
G\sum_jC_{ij}f_{j}(t),
\end{equation}
where $f_{j}(t)$ is a general term that describes some output node $j$ expressed to node $i$ which in most models is the dynamics of excitatory population at node $j$. Thus, if all nodes shared a common signal $\sigma(t)$ the dynamics at node $i$ will be modulated by this common factor and a local Gaussian noise term $N(0,1)$ (owing to noisy inputs in all the three models) expressed as:
\begin{eqnarray}
B_i(t) &=& \alpha N(0,1) + G\sum_jC_{ij}\sigma(t),\\
	   &=& \alpha N(0,1) + GD_{i}\sigma(t),
\end{eqnarray}
where $D_i$ is the node degree, and $\alpha$ determines the strength of the noise term which we set at $\alpha=1/2$. When we use the common signal as the mean signal of sub-10274, this noisy degree model in Fig~\ref{fig:NDM_G} can reproduce carpet plots and FC matrices similar to the preprocessing pipeline ICA-AROMA+2P with $G=4$. Therefore, we note that a model of the global signal modulated by node degree is roughly equivalent to highly synchronized models of rsfMRI, and is well correlated  to the balanced to the Balanced Excitation Inibition model ($r=0.99$). (think about how this fits in..)

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/NDMModel.png}
\caption{Wide spread deflections in the Noisy degree model. A B C D}\label{fig:NDM_G}
\end{figure*}

Secondly, in the weakly synchronized regime, there are fewer WSDs and there is closer match with that undergoing GSR or DiCER. However, within the experimental fMRI literature, a key argument against the use of GMR is that it may remove neurally related WSDs. Here, the presented models simulate globally coherent neuronal fluctuations, thus a key question is what remains post GSR. Following GMR, we see three types of artefactual effects in the three presented models in Fig.~\ref{fig:GMRModels}. For low $G$, WSDs are removed lowering VE1 and GFC (we exclude the noisy degree model as by construction, most of the signal is removed post GMR). This effect is present however, for all $G$ for the Hopf bifurcation model. Secondly, When there is a globally coherent signal that is delayed between nodes - such as those imposed by a travelling wave (Roberts et al.) - the mean signal can does not capture the WSD such as in the neural mass model for $G=3.5$ as in Fig.~\ref{fig:BalancedEI_G}~F thus leaving WSDs post GMR in Figure.~\ref{fig:GMRModels}~E,F. And finally, a residual artefactual signal can be left behind following GMR for the balanced EI models for $G=3$. In all models and global parameter variations, only the neural mass model for $G=2-3$ and the balaneced EI model for $G=2,4$ leaves residual structure in the FC matrix, indicating residual modulations above globally coherent ones. These are visual comparisons provide indications of the artefacts and FC matrix structure, quantiative comparisons with rsfMRI are described in the next section.

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/GMRModels.png}
\caption{GMR models, talk about what is happening. }\label{fig:GMRModels}
\end{figure*}

\subsection*{WSDs affect model fits}
With a characterization of the wide spread deflections in rsfMRI data and modelling, we can now explore how well these features correspond to quantitative model fits. In Figure~\ref{fig:FC_fits} we show the results of fitting functional connectivity similarity in the four models presented over three de-noising pipelines in the UCLA cohort. We note in all four models, model fits with regard to FC similarity are maximised with ICA-AROMA, with a global coupling factor that corresponds to model predictions with large global synchrony (as described above). This globally coherent feature is emphasized when comparing the FC of the model at optimal fits with the group average FC matrix in Fig.~\ref{fig:FC_fits}~E, i.e. regions are highly correlated to one another. The slight exception is the neural mass model, that has a lower overall fit. It is worth noting that the noisy degree model on data with large WSDs well in this metric and only contains one parameter ($G$).

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/FC_comparisons_all.png}
\caption{GMR models}\label{fig:FC_fits}
\end{figure*}

As decribed above, GMR applied rsfMRI models can reduce coherent signals thus an important question remains - what neuronal information is left post GMR and how well does this fit rsfMRI that has undergone reduction of WSDs with GMR or DiCER. When comparing rsfMRI models that have undergone GMR, we find that the performance of three of the presented models drop from $0.55$ to $3.55$ in the Balanced EI model, from $0.6$ to $0.2$ in the Hopf Bifurcation model, and $0.56$ to $0.1$ in the Noisy degree model to data that has undergone GMR or DiCER. In these models, WSDs dominate the dynamics thus their removal leaves residual dynamics that is less compatible with rsfMRI that has been de-noised past ICA-AROMA, which can be appreciated in the FC matrices in Figs.~\ref{fig:FC_fits}~F,G. Note that these results hold if the focus was restricted to unihemispheric dynamics. On the other hand, the dymamics in the neural mass model are weakly synchronized for $G<3.25$ and thus upon GMR the dynamics and the functional connectivity structure is unchanged - albeit shifted toward the negative, and the fitting of this model with the data is relatively unchanged.

Beyond static measures, dynamic measures of functional connectivity provide additional constraints on the models, thus the models can also be tested in terms of how well their dynamics match data. Here, we use phase-lagged coherence (what else do you call it Gustavo..) which, as stated in the methods, calculates the phase difference between all nodes for data within a window and correlates this with shifted versions of this window, as indicated in Fig.~\ref{fig:FCD_model_BEI}~A for subject 10274 in the UCLA cohort. From these figures, what can be appreciated is that the structure of the time-lagged coherence is reduced once the data has been de-noised in GMR or with DiCER. The distribution of the phase lagged coherence determines the dynamic nature of these fluctuations and Fig.~\ref{fig:FCD_model_BEI}~B indicates the aggregated histogram for all subjects. Here we emphasize that the fluctuations are correlated over time at different degrees - indicated by the spreading of the distribtution from 0 to 1. As the data is de-noised there is a shift to 0 indicating that this apparent time-lagged structure is largely dependent on the level of WSDs, thus WSDs bias these measures as well. When comparing the phase-lagged structure and histograms applied to the optimal models in the balanced EI model - the optimal model for blanced EI with ICA-AROMA and the optimal models for balanced EI+GSR for comparisons to ICA-AROMA+GSR and ICA-AROMA+DiCER. The shift toward 0 in the histogram is also present in the models that seem to generally match the histograms in the group data. To compare this quantiatively, a Komologrov-Smirnoff test (KS) is used to compare the histograms. The KS statistic from this test is called the Functional Connectivity Dynamics (FCD) and a small FCD implies a closer fit, and a large FCD (close to 1) implies a divergent fit. In Figure~\ref{fig:FCD_all_fits} this is calcualted in a similar fashion to Figs.~\ref{fig:FC_fits}.

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/FCD_BEI_data.png}
\caption{Phase - derived dynamic connectivity. Need to redo for longer and more sims.}\label{fig:FCD_model_BEI}
\end{figure*}

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/FCD_all.png}
\caption{Phase - derived dynamic connectivity. Need to redo for longer and more sims.}\label{fig:FCD_all_fits}
\end{figure*}

\subsection*{Targetting data with low WSDs: Heterogeniety}
Say after all, this it is better (and harder to tackle FCDs) this captures fluctations etc. This ANEC could be applied instead to the cleaner data and applied to all.

% Targetting the ANEC and discuss this.

% Here now say that WSDs greatly affect the results mainly because

% Also have to fix for BTF (need to run many simulations)

% Show Fit values, FCD values and DMN type of figures for all of them

% Then talk about this.

% Then talk about improving the models with heterongeous parameters on A -- doesn't work.

% Then talk about imrpovements with ANEC, it does work but at WHAT COST??!?!?

% Thats it.

% 3. How this effects both FCs and FCD show this for the optimal settings.

% Need to talk about the differences in the fittings on both FCD and FC, say that we previously showed how different coupling changes WSDs and then results change too.

% 3++ Influence on heterogenous parameters. - talk about ways to combat this.

% Three datasets show this - distribution of the G optimal values
% 4. Derivation of the Noisy-degree model, show on the B-EI model relative to GSR (in a figure with plots and things)

% we can derive this model - interesting relationship need to understand it - is it also related to ROI size? not sure.

% 5. Then talk about improvements -- ANEC can do so.

% +First show some carpet plots, talk about how these carpet plots show the level of coherent data, talk about how they represnet differenet levels of coherence and how they effect FC, dynamic FC.

% The problem with WSDs in data (and we dont correct for them) is that we then build and focus on models that also have large WSDs.

% Maybe then here we could show something with that on purpose.......

% We have simple measures there.

% We first have a summary of all the data, and how to interpret these (important points to highlight).

% Before modelling - we urge modellers to understand these effects.

% Show the results + show carpet plots and stuff

% Say that WSDs influence all of it

+++ Heterogeneity, things that can be done.

+++ ANEC

\section*{Discussion}
Global signal - noise or what? the problems within models.

\subsection*{Evidence against using data without some denoising} 
Need to discuss that a common reason not to use GMR (or even other WSD reducing types of analysis) is that it will impact neuronally driven WSDs - a feature in many models (as seen above). However models may just be fitting this and nothing else if theya re not careful.

\subsection*{The Balloon model}

A common forward model in the literature is to use the Balloon-windkessl model to simulate forward model dynamics. Although its use is common, a number of studies has invalidated the use of the Balloon model - namely the non-existence of the venous balloon. Although, the Balloon model may capture non-linear properites of the BOLD response in resting state it is likely that this is not achieved. Therefore we advise against the use of this model.

\section*{Notes}
+ Parcellation discussion increase

>>>>

The results are structured as follows. Firstly, we investigate the spatiotemporal characteristics of the data that need to be modelled. We then characterize the spatiotemporal characteristics of the models themselves, and compare the gross features of the models with data. Following this, we  describe the accuracy of the fits to data in the static and in the dynamic regime. Lastly, we show that we can improve the models by employing heterogenity in the model populations and in the estimates of effective functional connectivity.

@@

Firstly, an important technique to analyze the quality of resting state fMRI scans is to visualize the spatiotemporal time series of scans. As shown in Figure~\ref{fig:rsfMRIexp}C, (MPP looks pretty similar, might include a noisier subject..) that following minimal preprocessing (MPP) a ``low-motion'' subject from the UCLA cohort (sub-10274) contains a wealth of wide-spread signals that are times tied to motion events as captured by the mean framewise displacement (FD) - highlighted by the blue rectangles. Additional wide-spread signals have also been tied to physiological events such as respriation \cite{power2017simple,power2018ridding} or heart rate variability \cite{birn2006separating}. A robust de-noising method - ICA-AROMA with regression of white and CSF signals (+2P) mitigates some of these events however often with these alogirithms that use spatial ICA we are left with residual WSDs \cite{power2018ridding,power2013steps} which can be appreciated in Fig.~\ref{fig:rsfMRIexp}D. The link between these WSDs with physiology and motion thus treats these characterises these particular WSDs as noise. The effect of this WSD is expressed in the functional connectivity matrix in Fig~\ref{fig:rsfMRIexp}G, where the WSD causes a FC matrix that is postively correlated for almost all edges. This is an important issue because if the FC matrix is used to compare between groups, then large motion or physiology drives the changes in the edges of the FC matrix rather than underlying neuronal processes. For these reasons, further de-noising is often employed to rid rsfMRI of these artefactual signals. As described in the methods GMR is regularly used to rid rsfMRI of these influences, and as seen in Fig~\ref{fig:rsfMRIexp}E, it can help mitigate these effects with the side effect of introducing negative correlations due to this zero-centering \cite{murphy2009impact,saad2012-gsr-neg} which can be appreciated in Fig~\ref{fig:rsfMRIexp}H and Fig\ref{fig:rsfMRIexpbiphasic}H. Using the mean of all voxels to remove WSDs can be problematic if the WSD is made of components that exist in multiple phases such as the subject in Fig.~\ref{fig:rsfMRIexpbiphasic} where the common underlying WSD is expressed in two clear phases - i.e. the top half has the opposite sign of the bottom half. As can be apprecated in Fig.\ref{fig:rsfMRIexpbiphasic}D vs Fig.\ref{fig:rsfMRIexpbiphasic}E, GMR does not remove this WSD adequately, leaving behind a spatiotemporal time series that still contains artefactal WSD. This characterization of the types of WSDs is explained in \cite{aquino2019identifying} and can be thought to be of three types: weak (not shown here), single-phase, and biphasic. 
\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/figure_example.png}
\caption{Wide spread signals in a single-phase WSD - subject sub-10274 in the UCLA cohort. The first two signals express the A. signal averaged across all gray matter and B The framewise displacement (FD). The following four panels are carpet plots for the C. Minimal preprocessing pipeline (MPP), D. ICA-AROMA with white matter and CSF regression (+2P) pipeline (following MPP) E. Grey matter regression (GMR) following ICA-AROMA+2P pipeline, and F. Diffuse cluster estimation and regression (DiCER) following ICA-AROMA+2P pipeline. The last three panels show the corresponding functional connectivity matrices for G. ICA-AROMA+2P, H. ICA-AROMA+2P+GMR, and I. ICA-AROMA+2P+DiCER. }\label{fig:rsfMRIexp}
\end{figure*}

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/figure_example2.png}
\caption{Wide spread signals within a biphasic subject - sub-10724 in the UCLA cohort. The first two signals express the A. signal averaged across all gray matter and B The framewise displacement (FD). The following four panels are carpet plots for the C. Minimal preprocessing pipeline (MPP), D. ICA-AROMA with white matter and CSF regression (+2P) pipeline (following MPP) E. Grey matter regression (GMR) following ICA-AROMA+2P pipeline, and F. Diffuse cluster estimation and regression (DiCER) following ICA-AROMA+2P pipeline. The last three panels show the corresponding functional connectivity matrices for G. ICA-AROMA+2P, H. ICA-AROMA+2P+GMR, and I. ICA-AROMA+2P+DiCER. }\label{fig:rsfMRIexpbiphasic}
\end{figure*}

For these reasons, a new techinque - DiCER - was developed to targets these weak diffuse signals \cite{aquino2019identifying} in order to remove WSDs. On application of DiCER to the ICA-AROMA+2P output form all subjects, the carpet plots in Figures \ref{fig:rsfMRIexp}F and \ref{fig:rsfMRIexpbiphasic}F show the reduction of WSDs, whilsting revealing FC matrices in Figures \ref{fig:rsfMRIexp}~I and \ref{fig:rsfMRIexpbiphasic}~I that have structure without zero centering the functional connectivity matrix. We note that undergoing network discovery, DiCER appears to improve statistical sensivity of neuronal components - or at worst - on par with GMR \cite{aquino2019identifying}. These type of WSDs are present across the three datasets described, and the removal of the WSDs by DiCER or GMR generally improves QC-FC measures (not sure how much detail, perhaps will show QC-FC for UCLA,BZ and HCP).

@@

The above results illustrate that the nature of coherent structures are largely seen as artefactual and the presence of these structures are removed with varying degrees of de-noising. The varying degress of WSDs are important to biophysical models of resting state fMRI, as the models not only have to capture the overall functional connectivity but the typical dynamics within data. As described in the methods, a key feature of all biophysical models is the global coupling parameter $G$, that changes the dynamics of the models from un-coupled dynamical nodes ($G=0$) through to completely synchronized dynamics ($G>4$ in the models considered here). Therefore, the tuning of the model and the dynamic regime will depend on the quality of the data with regards to the level of WSDs. This can be appreciated by analzying the WSDs in the biophysical predictions in a similar fashion to the experimental data.

Firstly, for the Balanced Excitation Inihbition model Fig.~\ref{fig:BalancedEI_G}~A shows that varying the global coupling parameter $G$ largely changes the dynamics of the model. If we compare the carpet plots of varying levels of the coupling parameter $G$ we see visually that the level of WSDs increases as $G$ increases and is reflected in the global structure in the functional connectivity matrices in Fig.~\ref{fig:BalancedEI_G}~B. The level of WSDs can be quantified by using heuristics, here we use the variance explained by the first PC (VE1) or the global mean of all correlations in the FC matrix (here we use the absolute value of the FC matrix) (GFC). As shown in Figs~\ref{fig:BalancedEI_G}~C, and ~\ref{fig:BalancedEI_G}~D we can see that like in the carpet plots and the FC matrix, VE1 and GCF increases as $G$ increases. When comparing these measures to subject 10274 in the UCLA data we find that the level of synchrony needed to match the data depends on the level of WSDs within the data that are dependendent on the preprocessing pipeline. Overall, these features are shown in the neural mass model in Fig.~\ref{fig:BalancedEI_G}~F--I as well as the Hopf bifurcation model in Fig.~\ref{fig:BalancedEI_G}J--M. Where the transition from uncoupled noisy nodes to a synchronous pattern increases gradually in the Balanced EI model (Fig.~\ref{fig:BalancedEI_G}C,D) and in the Hopf bifurcation model(~\ref{fig:BalancedEI_G}~L,M), and is a phase transition in the Neural mass model (Fig.~~\ref{fig:BalancedEI_G}~H,I).

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/Models_all.png}
\caption{Wide spread deflections in biophysical models of restings state fMRI. In the three models presented here we show spatiotemporal carpet plots and functional connectivity matrices as a function of the global coupling parameter $G$ in (A,B), (F,G) and (J,K) for the balanced EI, neural mass and Hopf bifurcation models respectively. Quantatitive metrics of variance explained by the first PC (VE1) and global functional connectivity are shown in (C,D), (H,I), and (L,M) for the balanced EI, neural mass and Hopf bifurcation models respectively. In the metrics (i.e. the panels on the right) we calcualted these values for subject 10274 under three different preprocessing pipelines.}\label{fig:BalancedEI_G}
\end{figure*}

The above results indciate that data with WSDs requires models to be biased toward the synchronized regime, whereas data with reduced WSDs push data to the unsynchronized regime (i.e. lower $G$). We stress that the data used for comparison comes from the same subject here but undergoing different degrees of de-noising. We can explore this issue in the two regimes. Firstly, models within the synchronized regime appear to have a globaly coherent signal visually expressed by the large vertical stripes in Figs~\ref{fig:BalancedEI_G}~A,F,J with $G=3,3.25,2--8$ respecively. If we were to model this behaviour explicitly, we first note that in all three models the local dynamics at node $i$ are modulated by the influences owing to the sum (or mean) of the neighbouring nodes weighted by $C$ i.e. by a general factor
\begin{equation}
G\sum_jC_{ij}f_{j}(t),
\end{equation}
where $f_{j}(t)$ is a generic function that describes some output node $j$ expressed to node $i$. In most models $f_{j}(t)$ are the dynamics of excitatory population at node $j$ projected to node $i$. Thus, if all nodes shared a common signal $\sigma(t)$ the dynamics at node $i$ will be modulated by this common factor and a local Gaussian noise term $N(0,1)$ (owing to noisy inputs in all the three models) expressed as:
\begin{eqnarray}
z_i(t) &=& \alpha N(0,1) + G\sum_jC_{ij}\sigma(t),\\
	   &=& \alpha N(0,1) + GD_{i}\sigma(t),
\end{eqnarray}
where $D_i$ is the node degree, and $\alpha$ determines the strength of the noise term which we set at $\alpha=1/2$ when considering z-scored common signals $\sigma(t)$. Here, we describe neural dynamics however, since this is a linear process then we replace $z_i(t)$ with the estimate of the BOLD signal $B_i(t)$ if $\sigma(t)$ is within the same time scales. When we use the common signal as the mean signal of sub-10274, this noisy degree model in Fig~\ref{fig:NDM_G} can reproduce carpet plots and FC matrices similar to the preprocessing pipeline ICA-AROMA+2P with $G=4$. Therefore, a model of the global signal modulated by node degree is roughly equivalent to highly synchronized models of rsfMRI, and is well correlated to dynamics of the balanced to the Balanced Excitation Inibition model ($r=0.99$). (think about how this fits in..)

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/NDMModel.png}
\caption{Wide spread deflections in the noisy degree model for resting state fMRI. Here we show spatiotemporal carpet plots in A and functional connectivity matrices as a function of the global coupling parameter $G$ in B. Quantatitive metrics of variance explained by the first PC (VE1) are shown in C and global functional connectivity are shown in D. In the metrics (i.e. the C, and D) we calculated these values for subject 10274 under three different preprocessing pipelines.}\label{fig:NDM_G}
\end{figure*}

In the weakly synchronized regime, there are fewer WSDs and there is closer match with data undergoing GSR or DiCER. However, within the experimental fMRI literature, a key argument against the use of GMR is that it may remove neurally related WSDs. Here, the presented models simulate globally coherent neuronal fluctuations, thus a key question is what remains post GMR. Following GMR, we see three types of artefactual effects in the three presented models in Fig.~\ref{fig:GMRModels}. For low $G$, WSDs are removed lowering VE1 and GFC (we exclude the noisy degree model as by construction, as all of the non-noisy signal is removed post GMR). This effect is present however, for all $G$ for the Hopf bifurcation model. Secondly, When there is a globally coherent signal that is delayed between nodes - such as those imposed by a travelling wave (Roberts et al.) - the mean signal can does not capture the WSD such as in the neural mass model for $G=3.5$ as in Fig.~\ref{fig:BalancedEI_G}~F thus leaving WSDs post GMR in Figure.~\ref{fig:GMRModels}~E,F. 
% And finally, a residual artefactual signal can be left behind following GMR for the balanced EI models for $G=3$. 
In all models and global parameter variations, only the neural mass model for $G=2-3$ and the balanced EI model for $G=2,4$ leaves residual structure in the FC matrix, indicating residual modulations above globally coherent ones. These are visual comparisons and provide indications of the artefacts and FC matrix structure. Quantiative comparisons with rsfMRI are described in the next section.

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/GMRModels.png}
\caption{What is left post GMR in biophysical models of resting state fMRI. In the three models presented here we show spatiotemporal carpet plots and functional connectivity matrices following GMR as a function of the global coupling parameter $G$ in (A,B), (F,G) and (J,K) for the balanced EI, neural mass and Hopf bifurcation models respectively. Quantatitive metrics of variance explained by the first PC (VE1) and global functional connectivity are shown in (C,D), (H,I), and (L,M) for the balanced EI, neural mass and Hopf bifurcation models respectively. In the metrics (i.e. the panels on the right) we calcualted these values for subject 10274 under three different preprocessing pipelines.}\label{fig:GMRModels}
\end{figure*}

\subsection*{WSDs affect model fits}
With a characterization of the wide spread deflections in rsfMRI data and modelling, we can now explore how well these features correspond to quantitative model fits. In Figure~\ref{fig:FC_fits} we show the results of fitting functional connectivity similarity in the four models presented over three de-noising pipelines in the UCLA cohort. We note in all four models, model fits with regard to FC similarity are maximised with ICA-AROMA, with a global coupling factor that corresponds to model predictions with large global synchrony (as described above). This globally coherent feature is emphasized when comparing the FC of the model at optimal fits with the group average FC matrix in Fig.~\ref{fig:FC_fits}~E, i.e. regions are highly correlated to one another. The slight exception is the neural mass model, that has a lower overall fit. It is worth noting that the noisy degree model on data with large WSDs fits well with this analysis metric and only contains one parameter ($G$).

\begin{figure*}[ht!]
\includegraphics[width=1\textwidth]{figs/FC_comparisons_all.png}
\caption{Fitting of biophysical models to data. For the balanced EI model in A, the top panel shows the FC similarity of the model vs a group-averaged functional connectivity matrix under three different pipelines. The similiarity is compared with the model vs ICA-AROMA, model vs ICA-AROMA+GMR, model+GMR vs ICA-AROMA+GMR, model vs ICA-AROMA+DiCER and model+GMR vs ICA-AROMA+DiCER. Below this comparison, we show the optimal functional connectivity matrix when using the model vs ICA-AROMA, then the model+GMR vs ICA-AROMA+GMR and finally model+GMR vs ICA-AROMA+DiCER. In the three bottom plots the upper triangle is the model fits, and the lower triangle is the FC from the optimized fit. This is repeated for the Neural mass model in B, the Hopf bifurcation model in C and the noisy degree model in D.}\label{fig:FC_fits}
\end{figure*}

As decribed above, GMR applied to rsfMRI models can reduce coherent signals and thus an important question remains - what neuronal information is left post GMR and how well does this fit rsfMRI that has undergone a reduction of WSDs with GMR or DiCER. When comparing rsfMRI models that have undergone GMR, we find that the performance of three of the presented models drop from $0.55$ to $0.35$ in the Balanced EI model, from $0.6$ to $0.2$ in the Hopf Bifurcation model, and $0.56$ to $0.1$ in the Noisy degree model to data that has undergone GMR or DiCER. In these models, WSDs dominate the dynamics thus their removal leaves residual dynamics that is less compatible with rsfMRI that has been de-noised past ICA-AROMA, which can be appreciated in the FC matrices in Figs.~\ref{fig:FC_fits}~F,G. Note that these results hold if the focus was restricted to unihemispheric dynamics. On the other hand, the dymamics in the neural mass model are weakly synchronized for $G<3.25$ and thus upon GMR the dynamics and the functional connectivity structure is unchanged - albeit shifted toward the negative, and the fitting of this model with the data is relatively unchanged, however poor ($0.3$).

Beyond static measures, dynamic measures of functional connectivity provide additional constraints on the models, thus the models can also be tested in terms of how well their dynamics match data. Here, we use phase-lagged coherence (what else do you call it Gustavo..) which, as stated in the methods, calculates the phase difference between all nodes for data within a window and correlates this with shifted versions of this window, as indicated in Fig.~\ref{fig:FCD_model_BEI}~A for subject 10274 in the UCLA cohort. The phase-lagged coherence at $\phi_{ij}$ represents the difference between the synchronization level of the entire network at a time shift $\tau_i$ with a time shift $\tau_j$. Structure within $\phi_{ij}$ at different time shifts is usually interpreted as the dynamics of the persistence and decay of states of the network (i.e. at clusters we have similar coherence structure). What can be appreciated is that the structure of $\phi_{ij}$ is reduced once the data has been de-noised in GMR or with DiCER - i.e. the persistence of states are reduced, leaving only the diagonal as the main structure. Since the structure of $\phi_{ij}$ is dependent on a particular instance of the rsfMRI or model simulation, we instead focus on the distribution of the upper triangle of $\phi_{ij}$ instead as this determines the overall nature of these fluctuations. Figure~\ref{fig:FCD_model_BEI}~B indicates the aggregated histogram for all subjects. Here we emphasize that the fluctuations are correlated over time at different degrees - indicated by the spreading of the distribtution from 0 (i.e. uncorrelated structure over time) to 1 (instances where the structure is preserved over time shifts). As the data is de-noised there is a shift to 0 indicating that this apparent time-lagged structure is largely dependent on the level of WSDs, thus WSDs bias these measures as well. When comparing the phase-lagged structure and histograms applied to the optimal models in the balanced EI model - i.e. the optimal model for blanced EI with ICA-AROMA and the optimal models for balanced EI+GSR for comparisons to ICA-AROMA+GSR and ICA-AROMA+DiCER. The shift toward 0 in the histogram is also present in the models that seem to generally match the histograms in the group data. To compare this quantiatively, a Komologrov-Smirnoff test (KS) is used to compare the histograms for similarity. The KS statistic from this test is called the Functional Connectivity Dynamic (FCD) and a small FCD implies a closer fit, and a large FCD (close to 1) implies a divergent fit.
